### 2024-03-15 thesis/input_files/probabilistic.tex
March 15th 2024

\section{Kernel methods}
\subsection{Kernel quantile regression}
\subsection{Kernel herding}

>>>>

\section{Kernel methods}
\subsection{Kernel quantile regression}
The idea of quantile regression has been extended to kernel methods by Takeuchi et al. \cite{takeuchi2006nonparametric}.
There they minimize a risk plus regularizer defined as follows.
\begin{equation}\label{eq:kqr_min1}
    R[f]:=\frac{1}{m}\sum\limits_i=1^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|g\|_H^2
\end{equation}
where $f=g+b$, $g \in H$ and $b \in R$.
Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle+b$. Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
\begin{equation}\label{eq:kqr_min2}
    \begin{aligned}
    \min_{w,b} \quad & C \sum \limits_{i=1}^{m}
    (y_i-\langle w, \phi(x) \rangle-b)+ (1-q)(-y_i+\langle w, \phi(x) \rangle+b)+ \frac{1}{2}\|w\|^2\\
    \end{aligned}
    \end{equation}
Note the division by $\lambda$ so that $C=\frac{1}{\lambda m}$
We can next rephrase the optimisation in \ref{eq:kqr_min2} by introducing the slack variables $\xi_i$ and $\xi_i^*$.
\begin{equation}\label{eq:kqr_min3}
    \begin{aligned}
        \min_{w,b\xi_i,\xi_i^*} \quad & C \sum \limits_{i=1}^{m}
        q \xi_i+ (1-q)\xi_i^*+ \frac{1}{2}\|w\|^2\\
    \textrm{s.t.} \quad & y_i-\langle w, \phi(x) \rangle+b \leq \xi_i\\
    & -y_i+\langle w, \phi(x) \rangle+b \leq \xi_i^*\\
      &\xi_i\geq0    \\
      &\xi_i^*\geq0    \\
    \end{aligned}
    \end{equation}
In order to make it more compact, we rewrite equation \ref{eq:kqr_min3} in matrix notation.
\begin{equation}\label{eq:kqr_min4}
    \begin{aligned}
        \min_{w,b\xi,\xi^*} \quad & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w\\
    \textrm{s.t.} \quad & y-\Phi^\intercal w -b \preceq \xi\\
    & -y+\Phi^\intercal w +b \preceq \xi^*\\
      &\xi\succeq0    \\
      &\xi^*\succeq0    \\
    \end{aligned}
    \end{equation}
Consider now the associated lagrangian $L$ to \ref{eq:kqr_min4}
\begin{equation}\label{eq:kqr_min5}
    \begin{aligned}
    L(w,b,\xi,\xi^*)= & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w- \alpha^\intercal(\xi - y+\Phi^\intercal w +b)
    - (\alpha^*)^\intercal(\xi^* -y+\Phi^\intercal w +b)
    \\
    & -\nu^\intercal \xi - (\nu^*)^\intercal \xi^*
\end{aligned}
\end{equation}
The next step is deriving its dual formulation, since it is easier and more efficient to solve. This because the dual problem has the useful property of being always convex.

- dual problem
\begin{definition}
    The dual function associated to the lagrangia $L(x,\lambda)$ is given by $g(v)=\inf_x L(x,\lambda)$
\end{definition}
where $\lambda$ is called the lagrange multiplier of the optimization problem. Such dual formulation has an useful property, that is \begin{equation}\label{weak_duality}
    g(\lambda)\leq p^*
\end{equation}
where $p^*$ is the optimal value of your optimization problem.
Consider a simple lagrangian $L(x,\lambda)=f(x)+\sum \lambda_i i(x) +\sum v_i h(x)$, where $i(x)$ are inequality constraints while $e(x)$ are equality constraints of the problem. What we can be noted is that the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$. Therefore, the idea is that by maximizing the dual function subject to the contraint $\lambda \succeq 0$ we can obtain a approximate or perfect solution to the primal problem.
\\
To explain why we may or not be able to attain the best solution to the primal problem by maximizing its dual we have to introduce the concept of duality.
\\
We use $d^*$ to denote the optimal value of the lagrange dual problem; we can think of it as the best lower bound on $p^*$. The inequality \ref{weak_duality} is called weak duality. The difference $p^*-d^*$ is said the optimal duality gap; it is the gap between the optimal value of the primal problem and the best lower bound on it that can be obtained from the Lagrange dual function. Moreover, note that the optimal duality gap is always nonnegative.
\\
We say that strong duality holds, when the optimal duality gap is zero; in other words, the lagrange dual bound is tight.

- Slater's condition
- Apply Slater's theorem=>we have strong duality
- calculations
\subsection{Kernel herding}

### 2024-03-15 thesis/input_files/qr.tex
March 15th 2024

Quantile regression can be interpreted as an extension of standard regression. In this setting, you basically slice the dependent variables into quantiles and then fit a regression for each quantile. With standard regression we build a model for the conditional mean, conversely, with quantile regression we model the conditional quantile function for any desired quantile. 
Therefore, with quantile regression we are able to study the impact of covariates on quantiles directly.
\begin{definition}
    For any real valued random variable Y, we define its associated quantile function.
    \begin{equation}
        Q(q)=\inf{y:F(y)\geq q}
    \end{equation}
\end{definition}
Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solution to a simple optimization problem.
\\
For any $0<q<1$ consider the pinball loss function \ref{pinball} $\rho_q(u)=u(q-\mathbb{1}_{u<0})$. 
Such loss is minimized by the quantiles $Q(q)$.
We can estimate them by minimizing the expectation of $\rho_q(Y-g(x,\beta))$ with respect to the parameter $\beta$.
\\
Note, that in the special case $q=\frac{1}{2}$,  quantile regression corresponds exactly to standard regression with an absolute value loss function.
\\
It follows that the conditional linear quantile function $Q_Y(q|X=x)=x_i\beta(q)$ can be estimated by solving
\begin{equation}\label{}
    \hat{\beta(q)}=\argmin{\beta}\sum \rho_q   (y_i-x \beta)
\end{equation}
\\
Notice that, the cost function is not differentiable, therefore there is no analytical solution to the quantile regression problem. Nevertheless, we can easily solve it by employing linear programming.
\\
Furthermore, can extend to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above equation.
\\
In order to illustrate the approach, let us consider applying quantile regression to the maximum daily temperature dataset in Melbourne \cite{hyndman1996estimating}; models considered will be linear, gradient boosting machine, quantile forest\cite{} and kernel quantile regression.

- se si puo fare plot con stessi dati ma differenti f nel quantile regression

>>>>

Quantile regression can be interpreted as an extension of standard regression. In this setting, you basically slice the dependent variables into quantiles and then fit a regression for each quantile. With standard regression we build a model for the conditional mean, conversely, with quantile regression we model the conditional quantile function for any desired quantile. 
Therefore, with quantile regression we are able to study the impact of covariates on quantiles directly.
\begin{definition}
    For any real valued random variable Y, we define its associated quantile function.
    \begin{equation}
        Q(q)=\inf{y:F(y)\geq q}
    \end{equation}
\end{definition}
Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solution to a simple optimization problem.
\\
For any $0<q<1$ consider the pinball loss function \ref{pinball} $\rho_q(u)=u(q-\mathbb{1}_{u<0})$. 
Such loss is minimized by the quantiles $Q(q)$.
We can estimate them by minimizing the expectation of $\rho_q(Y-g(x,\beta))$ with respect to the parameter $\beta$.
\\
Note, that in the special case $q=\frac{1}{2}$,  quantile regression corresponds exactly to standard regression with an absolute value loss function.
\\
It follows that the conditional linear quantile function $Q_Y(q|X=x)=x_i\beta(q)$ can be estimated by solving
\begin{equation}\label{}
    \hat{\beta(q)}=\argmin{\beta}\sum \rho_q   (y_i-x \beta)
\end{equation}
\\
Notice that, the cost function is not differentiable, therefore there is no analytical solution to the quantile regression problem. Nevertheless, we can easily solve it by employing linear programming and convex optimisation \cite{boyd2004convex}.
\\
Furthermore, can extend to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above equation.
\\
In order to illustrate the approach, let us consider applying quantile regression to the maximum daily temperature dataset in Melbourne \cite{hyndman1996estimating}; models considered will be linear, gradient boosting machine, quantile forest\cite{meinshausen2006quantile} and kernel quantile regression.
- se si puo fare plot con stessi dati ma differenti f nel quantile regression
